---
title: "Data munging and figure preparation"
author: "Shyam Saladi (saladi@caltech.edu)"
date: "09/20/2016"
output: html_document
---

## Load libraries
Load all libraries necessary for subsequent data munging and plotting.
Some defaults are set here to make things easier throughout.
```{r libraries}
library(datamart) # `uconv` for unit conversion

library(magrittr)
library(tidyverse)
library(forcats)
library(dplyrExtras)

library(pROC)
library(ggbeeswarm)
library(cowplot)
library(viridis)
library(RColorBrewer)
theme_set(theme_cowplot(font_size = 7L))

library(parallel)
# currently provisioned resources
# cl <- c("brnode05", "brnode07", "brnode08") %>%
#     rep(20) %>%
#     c(rep("brnode11", 44)) %>%
#     makeCluster()

library(foreach)
library(doMC)
registerDoMC(cores = 20L)
#
# library(doParallel)
# registerDoParallel(cl)

# devtools::install("myutils")
library(myutils)
```

### Load NYCOMPS data for prediction function
```{r}
nycomps_env <- new.env()
load("large-scale.RData", nycomps_env)
```

### Define auc function for the subsequent calculations
```{r}
auc_w_ci <- partial(ci.auc, direction = "<",
                    method = "delong"#,
                    # boot.n = 1000,
                    # progress = "none",
                    # parallel = TRUE
                    )
```

# Calculate Pfam overlap between E. coli and NYCOMPS
```{r pfam_nycomps_overlap}
ecoli_pfam <- c("training/Daley_gfp.fna.faa.pfamA",
                "training/Daley_phoa.fna.faa.pfamA") %>%
    lapply(read.table, comment.char = '#', as.is = TRUE) %>%
    bind_rows %>%
    set_colnames(c('id', 'alignment_start', 'alignment_end',
                   'envelope_start', 'envelope_end', 'hmm_acc',
                   'hmm_name', 'type', 'hmm_start', 'hmm_end',
                   'hmm_length', 'bit_score', 'E-value',
                   'significance', 'clan')) %>%
    separate(id, "id")

ecoli_pfam_summary <- ecoli_pfam %>%
    group_by(id) %>%
    arrange(id, envelope_start) %>%
    summarize(hmm_arch = paste(hmm_acc, collapse = " "),
              clan_arch = paste(clan, collapse = " "))

nycomps_pfam <- c("large-scale/nycomps_only_pos.fna.faa.pfamA",
                  "large-scale/nycomps_only_neg.fna.faa.pfamA",
                  "large-scale/nycomps_mixed.fna.faa.pfamA") %>%
    lapply(read.table, comment.char = '#', as.is = TRUE) %>%
    bind_rows %>%
    as_data_frame %>%
    set_colnames(colnames(ecoli_pfam)) %>%
    separate(id, c("temp", "id"), convert = TRUE) %>%
    select(-temp) %>%
    # only keep Pfam families
    filter(type == "Family")

hmm_names <- nycomps_pfam %>%
    distinct(hmm_acc, hmm_name)
    
nycomps_pfam %<>%
    select(id, hmm_acc, clan) %>%
    # Check if the family is in the training set
    mutate(hmm_training = hmm_acc %in% ecoli_pfam$hmm_acc,
           clan_training = clan %in% ecoli_pfam$clan) %>%
    gather(family_type, family, hmm_acc, clan) %>%
    gather(training_type, in_training, hmm_training, clan_training) %>%
    filter((family_type == "hmm_acc" & training_type == "hmm_training") |
           (family_type == "clan" & training_type == "clan_training"),
           family != "No_clan") %>%
    select(-training_type) %>%
    distinct()
    
```

# merge pfam information with outcomes
```{r}
plasmid_names <- setNames(c("all",
                            "N", "C",
                            "C_LDAO", "MSGC.24",
                            "MSGC.28", "MSGC.7",
                            "MSGC.9", "MSGC.9_LDAO"),
                          c("all",
                            "His-FLAG-TEV- (N)", "-TEV-His (C)",
                            "-TEV-His (C, LDAO)", "His-GST-TEV- (24)",
                            "-TEV-His (28)", "His-TEV- (7)",
                            "His-MBP-TEV-(9)", "His-MBP-TEV- (9, LDAO)"))

nycomps_pfam_outcomes <- nycomps_env$nycomps_all_outcomes %>%
    mutate(cut = "Trials") %>%
    bind_rows(mutate(., plasmid_name = factor("all"))) %>%
    bind_rows(nycomps_env$nycomps_gene_outcomes %>%
                  mutate(cut = "Genes")) %>%
    select(-attempts) %>%
    mutate(gene_outcome = as.numeric(gene_outcome),
           plasmid_name = factor(plasmid_name, levels = plasmid_names,
                                 labels = names(plasmid_names))) %>%
    select(-posmix_outcome, -score_nornass, -cterm) %>%
    gather(outcome_type, outcome, gene_outcome, plasmid_outcome) %>%
    select(-outcome_type) %>%
    filter(!is.na(outcome)) %>%
    mutate(cut = factor(cut),
           plasmid_name = factor(plasmid_name)) %>%
    left_join(nycomps_pfam, by = "id")
```

# calculate AUC for subsets
```{r}
#overall for each group
nycomps_overall_auc <- nycomps_pfam_outcomes %>%
    distinct(cut, plasmid_name, id, score, outcome, family_type, in_training) %>%
    group_by(cut, plasmid_name, family_type, in_training) %>%
    filter(n() > 1, count_uniq(outcome) > 1, count_uniq(score) > 1) %>%
    do(calc_auc_roc(., method = "auc"))

# split by category
nycomps_pfam_auc <- nycomps_pfam_outcomes %>%
    group_by(cut, plasmid_name, family, family_type, in_training) %>%
    filter(n() > 1, count_uniq(outcome) > 1, count_uniq(score) > 1) %>%
    do(calc_auc_roc(., method = "auc"))
```

# auc for families in training or not
```{r}
p_hmm_acc_xlabels <- nycomps_pfam_outcomes %>%
    filter(family_type == "hmm_acc") %>%
    distinct(family, in_training) %>%
    group_by(in_training) %>%
    summarize(hmm_count = n()) %>%
    add_row(in_training = "NONE", hmm_count = "") %>%
    mutate(label = "In training") %>%
    mutate_rows(in_training == "FALSE", label = "Not in training") %>%
    mutate_rows(in_training == "NONE", label = "No Pfam") %>%
    mutate_rows(in_training != "NONE",
                label = paste0(label, "\n(", hmm_count, ")")) %>%
    arrange(in_training == "TRUE", in_training == "FALSE") %$%
    setNames(label, in_training)

p_hmm_acc <- nycomps_overall_auc %>%
    filter(family_type == "hmm_acc" | is.na(family_type)) %>%
    ungroup %>%
    mutate(in_training = factor(in_training,
                                levels = c("TRUE", "FALSE", "NONE"))) %>%
    mutate_rows(is.na(family_type), in_training = "NONE") %>%
    mutate_rows(is.na(lower95), upper95 = auc, lower95 = auc) %>%
    mutate_rows(is.na(plasmid_name), plasmid_name = paste0("genes", outcome)) %>%
    mutate(plasmid_name = fct_relevel(plasmid_name, "genes3", "genes2", "all"))

format_pos_count <- function(df) {
    df %>%
        arrange(pos_count) %$%
        unlist(pos_count) %>%
        divide_by(df$count[[1]]) %>%
        multiply_by(100) %>%
        round() %>%
        paste(collapse = "%, ") %>%
        paste0(., "% / ", df$count[[1]])
}

p_hmm_acc <- p_hmm_acc %>%
    filter(plasmid_name == "all" | outcome > 1) %>%
    ggplot(aes(x = in_training)) +
    geom_hline(yintercept = 0.5, linetype = "dashed", color = "lightgrey") +
    geom_pointrange(aes(ymin = upper95, y = auc, ymax = lower95,
                        group = as.numeric(plasmid_name),
                        color = ifelse(plasmid_name == "all", NA, outcome)),
                    size = 0.5, fatten = 0.5,
                    position = position_dodge(width = 0.5)) + #,
                    #data = filter(p_hmm_acc, cut == "Trials")) +
    # geom_pointrange(aes(ymin = upper95, y = auc, ymax = lower95,
    #                     group = -outcome, color = outcome),
    #                 size = 0.5, fatten = 0.5,
    #                 position = position_dodge(width = 0.1),
    #                 data = filter(p_hmm_acc, cut == "Genes")) +
    # geom_pointrange(aes(ymin = upper95, y = auc, ymax = lower95), color = "black",
    #                 size = 0.5, fatten = 0.5,
    #                 data = filter(p_hmm_acc, cut == "Trials", plasmid_name == "all")) +
    geom_text(aes(y = 0.47 - (.025 * as.numeric(cut)), label = pos_count,
                  color = ifelse(cut == "Trials", NA, 3)),
              size = 5*5/14,
              data = p_hmm_acc %>%
                  filter(plasmid_name %in% c("genes3", "genes2", "all")) %>%
                  group_by(cut, in_training) %>%
                  do(tibble(pos_count = format_pos_count(.)))) +
    ylab("AUC") +
    scale_x_discrete(expand = c(0, 0),
                     labels = p_hmm_acc_xlabels,
                     limits = names(p_hmm_acc_xlabels) %>% rev) +
    scale_y_continuous(expand = c(0.01, 0),
                       breaks = seq(0, 1, .05),
                       labels = function(x) x*100) +
    scale_color_continuous(low = "#A4A5A5", high = "#A10F1C", limits = c(1, 3),
                           guide = FALSE, na.value = "black") +
    # necessary for the labels
    coord_cartesian(ylim = c(0.495, 0.75)) +
    theme(axis.title.x = element_blank(),
          axis.text.x = element_text(size = 6),
          # necessary for the labels
          plot.margin = unit(c(4, 0, 19, 4), "pt"))

p_hmm_acc %<>% ggplotGrob()
p_hmm_acc$layout$clip <- "off"

ggdraw(p_hmm_acc)
```


```{r}
nycomps_pfam_outcomes_auc <- nycomps_pfam_auc %>%
    ungroup %>%
    filter(count > 5, plasmid_name == "all" | cut == "Genes") %>%
    # for confidence groups, need to do this
    mutate_rows(is.na(lower95), upper95 = auc, lower95 = auc) %>%
    group_by(cut) %>% 
    mutate(cumedist = row_number(auc)/length(auc),
           conf_grp = ifelse(auc <= 0.5, 0,
                             ifelse(lower95 <= 0.5, 0.5, 1))) %>%
    left_join(hmm_names, by = c("family" = "hmm_acc"))
```

## Considering multiple comparison adjustments:
In the case here, we don't actually drop anything out in the process, so it
seems within the FCR framework adjustment is not necesary (i.e., R == m). See:
Benjamini, Y., and D. Yekutieli. “False Discovery Rate-Adjusted Multiple
Confidence Intervals for Selected Parameters.” Journal of the American
Statistical Association 100, no. 469 (2005): 71–81.
https://en.wikipedia.org/wiki/False_coverage_rate
http://stats.stackexchange.com/a/148273/62183

# ecdf nycomps all outcomes - min max vectors
```{r}
get_ecdf_labels <- function(df)
    df %>%
    group_by(conf_grp) %>%
    summarize(count = n()) %>%
    ungroup %>%
    mutate(proport = round(count / sum(count)*100) %>% paste0("%")) %>%
    add_row(proport = nrow(df), conf_grp = NA)

p_nycomps_summary <- nycomps_pfam_outcomes_auc %>%
    filter(family_type == "hmm_acc", cut == "Trials") %>%
    group_by(cut, outcome)

p_nycomps_summary <- ggplot(p_nycomps_summary, aes(x = auc, y = cumedist)) +
    geom_segment(aes(x = lower95, xend = upper95, yend = cumedist),
                 alpha = 0.25, size = 0.25) +
    geom_point(aes(color = conf_grp), size = 0.15) +
    geom_text(aes(x = 0.06,
                  y = 0.84 + ifelse(is.na(conf_grp), -.05, conf_grp/9),
                  label = proport, color = conf_grp),
              size = 5*5/14,
              data = p_nycomps_summary %>% do({get_ecdf_labels(.)})) +
    xlab("AUC") +
    ylab("Cumulative Density") +
    scale_x_continuous(limits = c(-.01, 1.01), expand = c(0, 0),
                       labels = function(x) x*100) +
    scale_y_continuous(limits = c(-.01, 1.01), expand = c(0, 0),
                       labels = function(x)
                           paste0(specify_decimal(x*100, 0L), "%")) +
    scale_color_viridis(end = 0.7, direction = -1, guide = "none") +
    theme(legend.title = element_blank(),
          legend.position = c(0.2, 0.8),
          axis.line.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.text.y = element_blank(),
          axis.title.y = element_blank()) #+
   # facet_wrap(~cut + outcome, scales = "free", ncol = 2)

p_nycomps_summary %<>% ggplotGrob

p_nycomps_summary$layout %<>%
    mutate_rows(name == "panel", clip = "off")

ggdraw(p_nycomps_summary)
```

# write out table
```{r}
nycomps_pfam_outcomes_auc %>%
    ungroup() %>%
    filter(plasmid_name == "all" | is.na(plasmid_name), !is.na(family)) %>%
    filter(family_type == "hmm_acc", cut == "Trials") %>%
    mutate_rows(is.na(hmm_name), hmm_name = "") %>%
    mutate(outcome = factor(outcome, levels = c("1", "2", "3"),
                            labels = c("", "Mixed", "All Positive"))) %>%
    arrange(cut, desc(auc)) %>%
    transmute(`Pfam/Clan ID` = family,
              `Description` = hmm_name,
              `Family in_Training Set?` = in_training,
              `Outcome_Category` = cut,
              `Counts_(Genes/Trials)` = count,
              `Positive_Threshold` = outcome,
              `Positive_Outcomes` = pos_count,
              `AUC` = auc,
              `Lower Bound_95% CI` = lower95,
              `Upper Bound_95% CI` = upper95,
              `Best IMProve_Score Threshold` = best_thresh,
              `Best_Specificity` = best_spec,
              `Best_Sensitivity` = best_sens) %>%
    write_csv("forward-families/Table S4 - Family Performances.csv")
```

```{r}
format_auc <- function(auc, lower95, upper95, k = 1L)
    paste0(specify_decimal(auc*100, k), " (",
           specify_decimal(lower95*100, k), "-",
           specify_decimal(upper95*100, k), ")")

format_cnt <- function(pos_count, count, k = 0L)
    paste0(specify_decimal(pos_count/count*100, k), "%/", count)
```

# single vector families selected
```{r}
family_labels <-
    c(PF05425.10 = 'Copper\nresistance\nprotein D\n(PF05425)',
      PF02667.11 = 'Short-chain\nfatty-acid\ntransporter\n(PF02667)')

p_single_auc <- nycomps_pfam_outcomes %>%
    filter(family %in% names(family_labels), plasmid_name == "all") %>%
    separate(family, "family") %>%
    group_by(family) %>%
    do(calc_auc_roc(., method = "auc")) %>%
    ungroup() %>%
    arrange(family, -outcome_percentile) %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0(auc, " (", lower95, "-", upper95, ")"),
           thresh_x = row_number(family) - .25)

p_single <- nycomps_pfam_outcomes %>%
    filter(family %in% names(family_labels), plasmid_name == "all") %>%
    ggplot(aes(x = family, y = score)) +
    geom_point(aes(color = outcome, group = outcome), size = 0.25,
               position = position_beeswarm(dodge.width = .5)) +
    geom_segment(aes(y = best_thresh, yend = best_thresh,
                     x = thresh_x, xend = thresh_x + .5), 
                 size = 0.5, linetype = "dashed",
                 data = p_single_auc) + 
    ylab(expression("IMProve score")) +
    scale_color_continuous(low = "#A4A5A5", high = "#A10F1C") +
    scale_x_discrete(expand = c(0, 0),
                     labels = family_labels) +
    coord_flip() +
    theme(legend.position = "None",
          axis.title.y = element_blank(),
          plot.margin = unit(c(0, 0, 0, 0), "pt"))

p_single %<>% ggplotGrob()

p_single$layout %<>%
    mutate_rows(name == "panel", clip = "off")

ggdraw(p_single)
```

```{r}
p_single_auc_counts <- p_single_auc %>%
    group_by(family) %>%
    do(label = format_pos_count(.)) %>%
    ungroup() %>%
    mutate(label = unlist(label))

p_single_roc <- nycomps_pfam_outcomes %>%
    filter(family %in% names(family_labels), plasmid_name == "all") %>%
    separate(family, "family") %>%
    group_by(family) %>%
    do(calc_auc_roc(., method = "roc")) %>%
    group_by(outcome, add = TRUE) %>%
    split_groups() %>%
    ggplot(aes(color = outcome_percentile)) +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities)) +
    geom_point(aes(x = 1 - best_spec, y = best_sens),
               color = "black", size = 0.5,
               data = p_single_auc) +
    geom_text(aes(x = 0.75, y = .25, label = ci), color = "#A10F1C",
              size = 5*5/14, data = p_single_auc) +
    geom_text(aes(x = 0.75, y = 0.34, label = label), color = "#A10F1C",
              size = 5*5/14, data = p_single_auc_counts) +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    scale_color_continuous(limits = c(0, 100),
                           low = "#A4A5A5", high = "#A10F1C") +
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x) x*100) +
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x) x*100) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          axis.title = element_text(size = 6),
          strip.background = element_blank(),
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          strip.switch.pad.wrap = unit(c(0, 0, 0, 0), "lines"),
          strip.switch.pad.grid = unit(c(0, 0, 0, 0), "lines")) +
    facet_wrap(~family, ncol = 1, scales = "free")

# p3_gpcr_roc <- ggplotGrob(p3_gpcr_roc)
# p3_gpcr_roc$grobs[[13]]$grobs[[1]]$children[[2]]$children[[1]]$gp$col <-
#     "#D95F02"

ggdraw(p_single_roc)
```

```{r}
f_families <- ggdraw() +
    draw_plot(p_hmm_acc, x = 0, y = .45, width = 0.5, height = .55) +
    draw_plot(p_nycomps_summary, x = 0.5, y = .5, width = 0.5, height = .5) +
    draw_plot(p_single, x = 0, y = 0, width = .75, height = .43) +
    draw_plot(p_single_roc, x = .7, y = 0, width = .32, height = .52) +
    draw_plot_label(label = LETTERS[1:4],
                x = c(0, 0.5, 0, 0.74),
                y = c(1, 1, .45, 0.49), size = 8)

f_families %<>% ggplotGrob

save_plot(filename = "plots/fig5_3panel.pdf",
          plot = f_families,
          base_width = uconv(115, "mm", "in", "Length"),
          base_height = 4)

ggdraw(f_families)
```

```{r session}
save.image('fig_families.RData')
sessionInfo()
```
