---
title: "Data munging and Figure 3 preparation - Small scale datasets"
author: "Shyam Saladi (saladi@caltech.edu)"
date: "09/20/2016"
output: html_document
---

## Load libraries
Load all libraries necessary for subsequent data munging and plotting.
Some defaults are set here to make things easier throughout.
```{r libraries}
# if TRUE, then data will be written to file
write_out = FALSE

library(reshape2)
library(tidyverse)
library(dplyrExtras)

library(pROC)

library(RColorBrewer)
library(ggbeeswarm)
library(datamart) # `uconv` for unit conversion
library(cowplot)
theme_set(theme_cowplot(font_size = 7))

library(foreach)
library(doMC)

# devtools::install("myutils")
library(myutils)
library(doMC)
registerDoMC(cores = 4L)

# if TRUE, then data will be written to file
write_out = TRUE
```

### Load model data for prediction function
```{r}
model_env <- new.env()
load("model.RData", model_env)
svmpredict <- function(feature_df)
    prediction_fn(feature_df, model_env$daley_preprocess_params,
                  model_env$svm_model)
```

## archaea
```{r archaea_data_prep, cache=TRUE}
feat_col_spec = cols(.default = col_double(), title = col_character())

archaea_outcomes <- read_csv("small-scale/archaea.fna.allstats.csv",
                             col_types = feat_col_spec) %>%
    separate(title, c("Protein"), convert = TRUE)

archaea_outcomes$score <- svmpredict(archaea_outcomes)

archaea_outcomes <- read_csv("small-scale/archaea_S1.outcomes.csv",
                             col_types = cols(.default = col_character(),
                                              Outcome = col_double()),
                             na = c("n.c.", "n.t.")) %>%
    left_join(archaea_outcomes %>% select(Protein, score), by = "Protein") %>%
    filter(Measurement %in% c('WB/Gel',
                              'Scanning Densitometry of SDS-PAGE (%)'),
           !is.na(Outcome))

# Key for Archaea Outcomes:
# "None", "WB-only", "1-10% of MP", ">10% of MP"
# "0", "1", "5", "10"

archaea_median <- archaea_outcomes %>%
    filter(Measurement != 'Scanning Densitometry of SDS-PAGE (%)') %>%
    select(score) %>% unlist %>% unique %>% median
```

## archaea plotting
```{r archaea_plotting}
p3_archaea_sdspage_errbars <- archaea_outcomes %>%
    filter(Measurement == 'Scanning Densitometry of SDS-PAGE (%)') %>%
    group_by(Protein, score) %>%
    filter(n() > 1) %>%
    summarize(count = n(),
              sds_max = max(Outcome),
              sds_min = min(Outcome))

p3_archaea_sdspage <- archaea_outcomes %>%
    filter(Measurement == 'Scanning Densitometry of SDS-PAGE (%)') %>%
    # Don't plot points for values shown by the "error bars"
    anti_join(p3_archaea_sdspage_errbars %>%
                  gather(type, Outcome, sds_max:sds_min)) %>%
    # Repeat of this one outcome in the middle of an error bar
    # spread them apart a tiny bit for visualization
    mutate_rows(Protein == "Hbor39700", Vector == "pET52b(+)", Outcome == 12,
                Outcome = Outcome + 0.2) %>%
    mutate_rows(Protein == "Hbor39700", Vector == "pTTQ18", Outcome == 12,
                Outcome = Outcome - 0.2) %>%
    ggplot() +
    geom_vline(xintercept = archaea_median - .015,
               linetype = "dashed", color = "black", size = 0.5) +
    geom_point(aes(x = score, y = Outcome),
               color = "#A10F1C", size = 3, shape = 95) + 
    geom_errorbar(aes(x = score, ymin = sds_min, ymax = sds_max),
                  color = "#A10F1C", size = 0.27, width = 0.04,
                  data = p3_archaea_sdspage_errbars) +
    scale_y_continuous(expand = c(0, 0), limits = c(0, 40)) +
    scale_x_continuous(expand = c(0, 0), limits = c(-3.2, -1)) +
    ylab("Percentage of\nMembrane Protein") +
    xlab(expression("SVM" ^ "rank"*~"score")) + 
    scale_color_manual(values = brewer.pal(3, "Dark2")) +
    theme(axis.title = element_blank(),
          plot.margin = unit(c(1, 0, 2, 0), "mm"))

p3_archaea_sdspage
```

```{r}
p3_archaea_western_data <- archaea_outcomes %>%
    filter(Measurement != 'Scanning Densitometry of SDS-PAGE (%)') %>%
    group_by(Protein, score) %>%
    summarise(no = -sum(Outcome == 0L),
              wb = sum(Outcome == 1L),
              sdspage = sum(Outcome > 1L)) %>%
    gather(group, count, no:sdspage) %>%
    filter(count != 0L) %>% 
    # adjustments for visualization so bars are distinct
    mutate_rows(Protein == "MJ1319", score = score - 0.005) %>%
    mutate_rows(Protein == "Ta0252", score = score + 0.005)

p3_archaea_western <- ggplot(p3_archaea_western_data) + 
    geom_vline(xintercept = archaea_median - .015,
               linetype = "dashed", color = "black", size = 0.5) +
    geom_hline(yintercept = 0, alpha = 0.5,
               linetype = "dashed", color = "darkgrey") +
    geom_bar(aes(x = score, y = count), fill = "#A4A5A5",
             stat = "identity", width = .02,
             data = filter(p3_archaea_western_data,
                           group == "no", Protein == "MJ1319")) + 
    geom_bar(aes(x = score, y = count), fill = "#A4A5A5",
             stat = "identity", width = .02,
             data = filter(p3_archaea_western_data,
                           group == "no", Protein != "MJ1319")) + 
    geom_bar(aes(x = score, y = count, fill = group), position = "stack",
             stat = "identity", width = .02,
             data = filter(p3_archaea_western_data, group != "no")) +
    annotate("text", x = -2.9, y = 5.6, label = "Archaeal Transporters",
             size = 7*5/14, fontface = "bold") + 
    xlab(expression("SVM" ^ "rank"*~"score")) + 
    ylab("Count") +
    scale_fill_manual(values = c("#A10F1C", "#AD665D"), na.value = "black") +
    scale_y_continuous(expand = c(0, 0), limits = c(-6.1, 6),
                       breaks = seq(-6, 6, 3)) +
    scale_x_continuous(expand = c(0, 0), limits = c(-3.2, -1)) +
    theme(legend.position = "none",
              axis.title = element_blank(),
              plot.margin = unit(c(0, 0, 0, 0), "lines"))

p3_archaea_western
```

```{r}
format_pos_count <- function(df) {
    df %>%
        arrange(pos_count) %>%
        select(pos_count) %>% unlist %>%
        paste(collapse = ",") %>%
        paste0("list(", ., " / ", df$count[[1]], ")")
}
```

```{r}
archaea_auc <- archaea_outcomes %>%
    filter(Measurement != "Scanning Densitometry of SDS-PAGE (%)") %>%
    mutate_rows(Outcome == 10L, Outcome = 5L) %>%
    calc_auc_roc(outcome_col = "Outcome",
                 grouping_col = "Measurement",
                 method = "auc") %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0(auc, "~(", lower95, "-", upper95, ")")) %>%
    mutate_rows(outcome_threshold == 5L, y = .16) %>%
    mutate_rows(outcome_threshold == 1L, y = .07)

p3_archaea_roc <- archaea_outcomes %>%
    filter(Measurement != "Scanning Densitometry of SDS-PAGE (%)") %>%
    mutate_rows(Outcome == 10L, Outcome = 5L) %>%
    calc_auc_roc(outcome_col = "Outcome",
                 grouping_col = "Measurement",
                 method = "roc") %>%
    ggplot() +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold))) +
    geom_text(aes(x = .65, y = y, label = ci,
                  color = factor(outcome_threshold)),
              size = 5*5/14, parse = TRUE, data = archaea_auc) +
    annotate(geom = "text", x = .65, y = .24,
             label = format_pos_count(archaea_auc),
             size = 5*5/14, parse = TRUE) + 
    xlab("False Positive Rate") +
    ylab("True Positive Rate") +
    ggtitle("Archaeal Transporters") + 
    scale_color_manual(values = setNames(c("#AD665D", "#A10F1C"), c(1L, 5L))) +
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          axis.title = element_text(size = 6))

p3_archaea_roc
```

### M. TUBERCULOSIS OUTCOMES AND SCORES (FROM TIM CROSS'S FIRST PAPER)
```{r mt_data_prep, cache = TRUE}
tcross_outcomes <- read_csv("small-scale/tcross.fna.allstats.csv",
                             col_types = feat_col_spec) %>%
    separate(title, c("ORF"))

tcross_outcomes$score <- svmpredict(tcross_outcomes)

tcross_outcomes <- read_csv("small-scale/tcross_mt_outcomes.csv") %>%
    mutate(cloned = Cloned,
           host = strain,
           IB_CB = IB_CB*2,
           SOL_CB = SOL_CB*2,
           MEM_CB = MEM_CB*2,
           IB_WB = as.numeric(IB_WB),
           SOL_WB = as.numeric(SOL_WB),
           MEM_WB = as.numeric(MEM_WB)) %>%
    mutate_rows(is.na(IB_WB), IB_WB = 0) %>%
    mutate_rows(is.na(SOL_WB), SOL_WB = 0) %>%
    mutate_rows(is.na(MEM_WB), MEM_WB = 0) %>%
    mutate(IB = pmax(IB_CB, IB_WB),
           SOL = pmax(SOL_CB, SOL_WB),
           MEM = pmax(MEM_CB, MEM_WB)) %>%
    select(ORF, cloned, host, IB, SOL, MEM) %>%
    gather(fraction, outcome, IB:MEM) %>%
    mutate(outcome = factor(outcome,
                            levels = c(0, 1, 2),
                            labels = c("No","WB","CB"),
                            ordered = TRUE),
           fraction = factor(fraction,
                             levels = c("MEM", "SOL", "IB"),
                             labels = c("MEM", "SOL", "IB"))) %>%
    left_join(tcross_outcomes %>% select(ORF, score), by = "ORF")

# > select(tcross_outcomes, ORF, cloned, host) %>% distinct %>% group_by(cloned, host) %>% summarize(tot_cloned = length(cloned))
# Source: local data frame [3 x 3]
# Groups: cloned [?]
# 
#   cloned              host tot_cloned
#    (chr)             (chr)      (int)
# 1      C CodonPlus-RP(DE3)         12
# 2      N          C42(DE3)          4
# 3      N CodonPlus-RP(DE3)         93

# > table(tcross_outcomes$outcome)
#  No  WB  CB 
# 208  66  53 
# > table(as.numeric(tcross_outcomes$outcome))
#   1   2   3 
# 208  66  53
```

## mt plotting
```{r mt_plotting}
p3_mt <- tcross_outcomes %>%
    filter(fraction == "MEM") %>%
    ggplot() +
    geom_vline(aes(xintercept = median(score)),
               linetype = "dashed", size = 0.5) + 
    geom_jitter(aes(x = score, y = outcome, color = as.numeric(outcome)),
                size = .95, position = position_quasirandom(varwidth = TRUE),
            data = filter(tcross_outcomes, fraction == "MEM")) +
    annotate("text", x = -3.7, y = 3.1, label = "M. tuberculosis",
             size = 7*5/14, fontface = "bold.italic") + 
    scale_y_discrete(labels = c("WB" = "Western\nBlot",
                                "CB" = "Coomassie\nBlue",
                                "Neither" = "No\nExpression")) +
    scale_x_continuous(expand = c(.01, 0), breaks = c(-5:5)) +
    scale_color_gradient(low = "#A4A5A5", high = "#A10F1C",
                         na.value = "black") +
    theme(axis.title = element_blank(),
          legend.position = "None",
          plot.margin = unit(c(0, 0, 0, 0), "lines"))

p3_mt
```

```{r}
tcross_auc <- tcross_outcomes %>%
    filter(fraction == "MEM") %>%
    mutate(outcome = as.numeric(outcome)) %>% 
    calc_auc_roc(grouping_col = "fraction",
                 method = "auc") %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0(auc, "~(", lower95, "-", upper95, ")")) %>%
    mutate_rows(outcome_threshold == 3L, y = .16) %>%
    mutate_rows(outcome_threshold == 2L, y = .07)

p3_mt_roc <- tcross_outcomes %>%
    filter(fraction == "MEM") %>%
    mutate(outcome = as.numeric(outcome)) %>% 
    calc_auc_roc(grouping_col = "fraction",
                 method = "roc") %>%
    ggplot() +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold))) +
    geom_text(aes(x = .65, y = y, label = ci,
                  color = factor(outcome_threshold)),
              size = 5*5/14, parse = TRUE, data = tcross_auc) +
    annotate(geom = "text", x = .65, y = .25,
             label = format_pos_count(tcross_auc),
             size = 5*5/14, parse = TRUE) +
    scale_color_manual(values = ggplot_build(p3_mt)$data[[2]] %>% 
                           arrange(desc(-group)) %>%
                           select(colour) %>%
                           unique %>% unlist %>% unname %>% tail(-1),
                       na.value = "black") +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    ggtitle("M. tuberculosis") +
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          plot.title = element_text(size = 7, face = "italic"),
          axis.title = element_text(size = 6))

p3_mt_roc
```

## GPCRs
```{r gpcr_data_prep, cache=TRUE}
lundstrom_outcomes <-
    read_csv("small-scale/lundstrom1.fna.allstats.csv",
             col_types = feat_col_spec)

lundstrom_outcomes$score <- svmpredict(lundstrom_outcomes)

lundstrom_outcomes <-
    read_csv("small-scale/lundstrom1.060915.outcomes.csv") %>%
    filter(trust == 1) %>%
    left_join(lundstrom_outcomes %>% select(title, score),
              by = c("new ids" = "title")) %>%
    gather(host, outcome, `E. coli`:SFV) %>%
    filter(outcome != "nd") %>%
    mutate(outcome = factor(outcome,
                            labels = c("No", "Low", "Medium", "High"),
                            levels = c("0", "1", "2", "3"),
                            ordered = TRUE))

lundstrom_median <- lundstrom_outcomes %>%
    group_by(host) %>%
    summarize(median = median(score))

# > table(lundstrom_outcomes$outcome)
#     No    Low Medium   High 
#     56     77     54     87 
# > table(as.numeric(lundstrom_outcomes$outcome))
#  1  2  3  4 
# 56 77 54 87
```

## GPCR plotting
We need to do it this way to break the x axis.
```{r gpcr_plotting}
p3_gpcr <- ggplot(filter(lundstrom_outcomes,
                         host %in% c("E. coli", "P. pastoris"))) +
    geom_quasirandom(aes(x = host, y = score, color = outcome),
                     dodge.width = .7, varwidth = TRUE, size = 1) +
    geom_segment(aes(y = median, yend = median, x = 1.5, xend = 2.5),
                 size = 0.5, linetype = "dashed", 
                 data = filter(lundstrom_median, host == "E. coli")) +
    geom_segment(aes(y = median, yend = median, x = 0.5, xend = 1.5),
                 size = 0.5, linetype = "dashed", 
                 data = filter(lundstrom_median, host == "P. pastoris")) +
    scale_color_manual(
        values = setNames(c('#A4A5A5', '#AC7B74', '#AA4F46', '#A10F1C'),
                          c("No", "Low", "Medium", "High"))) +
    scale_x_discrete(expand = c(0, 0)) +
    ylab(expression("SVM" ^ "rank"*~"score")) + 
    coord_flip() +
    theme(legend.position = "None")

p3_gpcr_sub_left <- p3_gpcr + 
    scale_y_continuous(expand = c(1, 0), breaks = c(-10)) +
    coord_flip(ylim = c(-12, -10))  + 
    theme(axis.title.x = element_blank(),
          axis.text.y = element_text(face = "italic",
                                     color = brewer.pal(3, "Dark2")[c(2, 1)]),
          axis.title.y = element_blank(),
          plot.margin = unit(c(2, 0, 2, 0), "mm"))

p3_gpcr_sub_right <- p3_gpcr + 
    annotate("text", y = -4.6, x = 2.36, label = "Mammalian GPCRs",
             size = 7*5/14, fontface = "bold") + 
    scale_y_continuous(limits = c(-5.4, 0.6), expand = c(0, 0)) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.line.y = element_blank(),
          plot.margin = unit(c(2, 0, 2, 0), "mm"))

# for some reason geom_quasirandom has to be added before geom_segment,
# so change the order to have the segments in back here:
p3_gpcr_sub_right <- ggplotGrob(p3_gpcr_sub_right)
p3_gpcr_sub_right$grobs[[4]]$childrenOrder <- 
    p3_gpcr_sub_right$grobs[[4]]$childrenOrder[c(1,3,4,2,5)]

p3_gpcr_main <- plot_grid(p3_gpcr_sub_left, p3_gpcr_sub_right,
                          align = "h", nrow = 1, rel_widths = c(1.3, 10))
ggdraw(p3_gpcr_main)
```

```{r}
lundstrom_auc <- lundstrom_outcomes %>%
    filter(host %in% c("E. coli", "P. pastoris")) %>%
    mutate(outcome = as.numeric(outcome)) %>%
    calc_auc_roc(grouping_col = "host") %>%
    arrange(group, -outcome_threshold) %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0("'", auc, "'~(", lower95, "-'", upper95, "')"),
           # Assigning values directly  is a bit precarious,
           # but it should be okay with the `arrange` above
           x = c(rep(.65, 3), rep(.75, 3)),
           y = rep(c(.25, .16, .07), 2))

lundstrom_auc_counts <- lundstrom_auc %>%
    group_by(group) %>%
    do(label = format_pos_count(.)) %>%
    ungroup %>%
    mutate(label = unlist(label)) %>%
    mutate_rows(group == "E. coli", x = .65) %>%
    mutate_rows(group == "P. pastoris", x = .75)

p3_gpcr_roc <- lundstrom_outcomes %>%
    filter(host %in% c("E. coli", "P. pastoris")) %>%
    mutate(outcome = as.numeric(outcome)) %>%
    calc_auc_roc(grouping_col = "host", method = "roc") %>%
    ggplot() +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold))) +
    geom_text(aes(x = x, y = y, label = ci,
                  color = factor(outcome_threshold)),
              parse = TRUE, size = 5*5/14, data = lundstrom_auc) +
    geom_text(aes(x = x, y = 0.34, label = label), parse = TRUE,
              size = 5*5/14, data = lundstrom_auc_counts) + 
    scale_color_manual(
        values = setNames(c('#AC7B74', '#AA4F46', '#A10F1C'),
                          c('2', '3', '4'))) +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    ggtitle("Mammalian GPCRs") + 
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x) x*100) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x) x*100) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          axis.title = element_text(size = 6),
          strip.background = element_blank(),
          strip.text = element_text(size = 6, face = "italic",
                                    color = brewer.pal(3, "Dark2")[1:2]),
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          strip.switch.pad.wrap = unit(c(0, 0, 0, 0), "lines"),
          strip.switch.pad.grid = unit(c(0, 0, 0, 0), "lines")) +
    facet_wrap(~group, ncol = 1, scales = "free")

p3_gpcr_roc <- ggplotGrob(p3_gpcr_roc)
p3_gpcr_roc$grobs$strip_t2$grobs[[1]]$children[[2]]$children[[1]]$gp$col <- 
    "#D95F02"

ggdraw(p3_gpcr_roc)
```

## fig3 main
```{r fig3_main_plotting}
p3_main <- ggdraw() +
    draw_plot(
        plot_grid(p3_archaea_western, p3_archaea_sdspage, p3_mt, NULL,
                  labels = c("a", "","","e"), label_size = 8, ncol = 1,
                  align = "v", rel_heights = c(1.8, 1.8, 2.2, 3.2)),
        x = 0, y = 0, width = .8, height = 1) +
    draw_plot_label("c", x = 0, y = .62, size = 8) +
    draw_plot(p3_gpcr_main, x = 0.003, y = 0, width = .797, height = .37) +
    annotate("segment", x = .0975, xend = .0935, y = .063, yend = .072) +
    annotate("segment", x = .1010, xend = .0970, y = .063, yend = .072) +
    draw_label("Number of Conditions",
               x = .03, y = .912, angle = 90, size = 7) +
    draw_label("Percentage of\nMembrane Protein",
               x = .03, y = .71, angle = 90, size = 7) +
    draw_label("Coomassie Blue", x = .22, y = .945,
               size = 6, colour = "#A10F1C") +
    draw_label("Western Blot", x = .22, y = .925,
               size = 6, colour = "#AD665D") +
    draw_label("No Expression", x = .22, y = .895,
               size = 6, colour = "#A4A5A5") +
    draw_plot(
        plot_grid(p3_archaea_roc, p3_mt_roc, NULL,
                  labels = c("b", "d", "f"), label_size = 8,
                  ncol = 1, align = "hv"),
        x = .82, y = .25, width = .16, height = .75) +
    draw_plot(p3_gpcr_roc, x = .815, y = 0, width = .17, height = .5)

if (write_out) {
    save_plot(filename = "plots/fig3_w_panels.pdf", plot = p3_main,
              base_width = uconv(183, "mm", "in", "Length"), base_height = 5)
}

p3_main
```

## hpylori
```{r hpylori_data_prep, cache = TRUE}
hpylori_outcomes <- read_csv("small-scale/hpylori_psakis.ffn.allstats.csv",
                             col_types = feat_col_spec)

hpylori_outcomes$score <- svmpredict(hpylori_outcomes)

hpylori_outcomes <- read_csv("small-scale/hpylori-psakis.csv",
                             col_types = cols(.default = col_character(),
                                              Outcome = col_integer()),
                             na = "x") %>%
    filter(!is.na(Outcome)) %>% 
    transmute(gene = toupper(Code),
              group = factor(Vector,
                          levels = c("pET", "LAC / A", "LAC / M",
                                     "TEV / A", "TEV / M"),
                          labels = c("pET28 / pET36", "pQL60Lac / Automatic",
                                     "pQL60Lac / Manual", 
                                     "pQTev / Automatic", "pQTev / Manual")),
              vector = factor(substring(Vector, 1, 3)),
              scoring = factor(substrRight(Vector, 1)),
              outcome = Outcome) %>%
    inner_join(hpylori_outcomes %>% select(title, score),
               by = c("gene" = "title")) %>%
    group_by(gene, vector, score) %>%
    # average multiple outcomes for given genes
    summarize(outcome = mean(outcome)) %>%
    ungroup

hpylori_median <- hpylori_outcomes %>%
    distinct(gene, score) %>%
    summarize(median = median(score)) %>%
    unlist(use.names = FALSE)
```

## hpylori plotting
```{r hpylori_plotting}
p3_hpylori <- ggplot(hpylori_outcomes) + 
    geom_vline(xintercept = hpylori_median,
               size = 0.5, linetype = "dashed") +
    geom_point(aes(x = score, y = outcome, color = outcome),
               size = 0.75, position = position_quasirandom(varwidth = TRUE)) +
    annotate("text", x = -6.5, y = 3.25, label = "H. pylori",
             size = 7*5/14, fontface = "bold.italic") + 
    scale_color_gradient(low = "#A4A5A5", high = "#A10F1C",
                         na.value = "black") +
    scale_x_continuous(breaks = c(-8:1), expand = c(0.01, 0.01)) +
    scale_y_continuous(breaks = seq(0, 3, 0.5), expand = c(0.02, 0.02)) +
    xlab(expression("SVM" ^ "rank"*~"score")) + 
    ylab("Reported Outcome") +
    theme(legend.position = "None",
              axis.title.y = element_blank(),
              axis.title.x = element_blank(),
              plot.margin = unit(c(2, 0, 0, 0), "mm"))

p3_hpylori
```


```{r}
p3_hpylori_roc <- hpylori_outcomes %>%
    mutate(group = "group") %>%
    calc_auc_roc(grouping_col = "group", method = "roc") %>%
    ggplot() +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold)), size = 0.25) +
    scale_color_manual(
        values = colorRampPalette(c("#A4A5A5", "#A10F1C"))(7)[2:7]) +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    ggtitle("H. pylori") + 
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) +
    theme(legend.position = "None", aspect.ratio = 1,
              plot.margin = unit(c(0, 0, 0, 0), "lines"),
              plot.title = element_text(size = 7, face = "italic"),
              axis.title = element_text(size = 6))

p3_hpylori_roc
```

```{r}
hpylori_auc <- hpylori_outcomes %>%
    mutate(group = "group") %>%
    calc_auc_roc(grouping_col = "group", method = "auc") %>%
    arrange(-outcome_threshold) %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0("'", auc, "'~('", lower95, "'-'", upper95, "')"),
           x = rep(.5, 6),
           y = c(.60, .50, .40, .30, .20, .10))

p3_hpylori_auc <- ggplot(hpylori_auc) + 
    geom_text(aes(x = x, y = y, label = ci,
                  color = factor(outcome_threshold)),
              parse = TRUE, size = 5*5/14) +
    annotate(geom = "text", x = .5, y = .70,
             label = format_pos_count(hpylori_auc),
             size = 5*5/14, parse = TRUE) +
    scale_color_manual(
        values = colorRampPalette(c("#A4A5A5", "#A10F1C"))(7)[2:7]) +
    scale_x_continuous(expand = c(0, 0), limits = c(0, 1)) +
    theme(legend.position = "None",
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          plot.title = element_blank(),
          axis.title = element_blank(),
          axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())

p3_hpylori_auc
```

## t maritima Koth
```{r koth_data_prep, cache=TRUE}
tmaritima_outcomes <-
    read_csv("small-scale/koth_tmaritima.ffn.allstats.csv",
             col_types = feat_col_spec) %>% 
    separate(title, c("Gene"))

tmaritima_outcomes$score <- svmpredict(tmaritima_outcomes)

tmaritima_outcomes <- read_csv("small-scale/koth_tmaritima.outcomes.csv") %>%
    inner_join(tmaritima_outcomes %>% select(Gene, score),
               by = c("Gene" = "Gene")) %>%
    mutate(Purified = Purified * 2,
           outcome = factor(pmax(Expressed, Purified),
                            labels = c("Neither", "Expressed", "Purified"),
                            levels = c(0, 1, 2),
                            ordered = TRUE))
```

### Koth plotting
```{r koth_plotting}
p3_tmaritima <- ggplot(tmaritima_outcomes) +
    geom_vline(aes(xintercept = median(score)), linetype = "dashed") +
    geom_quasirandom(aes(x = score, y = factor(outcome),
                         color = as.numeric(outcome)),
                     varwidth = TRUE, width = 0.3, size = 0.75) +
    annotate("text", x = -4.6, y = 3.36, label = "T. maritima",
             size = 7*5/14, fontface = "bold.italic") +
    scale_color_gradient(low = "#A4A5A5", high = "#A10F1C",
                         na.value = "black") +
    xlab(expression("SVM" ^ "rank"*~"score")) +
    theme(legend.position = "None",
              axis.title = element_blank())

p3_tmaritima
```

```{r}
koth_auc <- tmaritima_outcomes %>%
    mutate(outcome = as.numeric(outcome),
           group = "group") %>%
    calc_auc_roc(grouping_col = "group", method = "auc") %>%
    arrange(-outcome_threshold) %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0("'", auc, "'~('", lower95, "'-'", upper95, "')"),
           y = c(.16, .07))

p3_tmaritima_roc <- tmaritima_outcomes %>%
    mutate(outcome = as.numeric(outcome),
           group = "group") %>%
    calc_auc_roc(grouping_col = "group", method = "roc") %>%
    ggplot() +
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold))) +
    geom_text(aes(x = .65, y = y, label = ci,
                  color = factor(outcome_threshold)),
              parse = TRUE, size = 5*5/14, data = koth_auc) +
    annotate(geom = "text", x = .65, y = .24,
             label = format_pos_count(koth_auc),
             size = 5*5/14, parse = TRUE) +
    scale_color_manual(values = c("#AD665D", "#A10F1C"),
                       na.value = "black") +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    ggtitle("T. maritima") +
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          plot.title = element_text(size = 7, face = "italic"),
          axis.title = element_text(size = 6))

p3_tmaritima_roc
```

## surade
```{r surade_data_prep, cache=TRUE}
surade_outcomes <- read_csv("small-scale/surade.fna.allstats.csv",
                             col_types = feat_col_spec) %>%
    separate(title, c("GI number"))

surade_outcomes$score <- svmpredict(surade_outcomes)

surade_outcomes <- read_csv("small-scale/surade.060915.outcomes.csv",
                            col_types = cols(.default = col_character(),
                                             Outcome = col_integer()),
                            na = c("CF", "NT")) %>%
    inner_join(surade_outcomes %>% select(`GI number`, score),
               by = "GI number") %>%
    filter(!is.na(Outcome)) %>% 
    mutate(Vector = factor(Vector, 
                           labels = c("E. coli pTTQ18-A", "E. coli pTTQ18-C",
                                      "E. coli pQE-A", "E. coli pQE-C",
                                      "E. coli pBAD-A", "E. coli pBAD-C",
                                      "L. lactis pNZ-A", "L. lactis pNZ-C"),
                           levels = c("pTTQ18-A", "pTTQ18-C",
                                      "pQE-A", "pQE-C",
                                      "pBAD-A", "pBAD-C",
                                      "pNZ-A", "pNZ-C"))) %>%
    # Only consider, pET variety vectors in E. coli
    filter(Vector %in% c("E. coli pTTQ18-A", "E. coli pTTQ18-C",
                         "E. coli pQE-A", "E. coli pQE-C"))

surade_median <- surade_outcomes %>%
    distinct(score) %>%
    unlist %>% median
```

## surade plotting
```{r surade_plotting}
p3_surade <- ggplot(surade_outcomes) +
    geom_vline(xintercept = surade_median, linetype = "dashed") +
    geom_quasirandom(aes(x = score, y = factor(Outcome), color = Outcome),
                     varwidth = TRUE, width = 0.3, size = 0.75) + 
    annotate("text", x = -2.7, y = 3.36,
             label = "Microbial Secondary Transporters",
             size = 7*5/14, fontface = "bold") + 
    scale_x_continuous(breaks = seq(-4, 1, 0.5)) +
    scale_color_gradient(low = "#A4A5A5", high = "#A10F1C",
                         na.value = "black") +
    xlab(expression("SVM" ^ "rank"*~"score")) + 
    ylab("Dot Density") +
    theme(legend.position = "None",
          axis.title.y = element_blank())

p3_surade
```

```{r}
surade_auc <- surade_outcomes %>%
    mutate(group = "pETonly") %>%
    calc_auc_roc(grouping_col = "group", outcome_col = "Outcome",
                 method = "auc") %>%
    arrange(-outcome_threshold) %>%
    mutate(auc = specify_decimal(auc*100),
           lower95 = specify_decimal(lower95*100),
           upper95 = specify_decimal(upper95*100),
           ci = paste0("'", auc, "'~(", lower95, "-", upper95, ")")) %>%
    mutate_rows(outcome_threshold == 10, y = 0.07) %>%
    mutate_rows(outcome_threshold == 100, y = 0.16)

p3_surade_roc <- surade_outcomes %>%
    mutate(group = "pETonly") %>%
    calc_auc_roc(grouping_col = "group", outcome_col = "Outcome",
                 method = "roc") %>%
    ggplot() + 
    geom_abline(slope = 1, intercept = 0,
                linetype = "dashed", color = "black") +
    geom_path(aes(x = 1 - specificities, y = sensitivities,
                  color = factor(outcome_threshold))) +
    geom_text(aes(x = .65, y = y, label = ci, color = factor(outcome_threshold)),
              parse = TRUE, size = 5*5/14, data = surade_auc) +
    annotate(geom = "text", x = .65, y = .24,
             label = format_pos_count(surade_auc),
             size = 5*5/14, parse = TRUE) +
    scale_color_manual(
        values = colorRampPalette(c("#A4A5A5", "#A10F1C"))(3)[2:3],
        na.value = "black") +
    ylab("True Positive Rate") +
    xlab("False Positive Rate") +
    ggtitle("Microbial\nsecondary transporters") +
    scale_y_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) + 
    scale_x_continuous(expand = c(0, 0), breaks = c(0, .5, 1),
                       limits = c(0, 1), labels = function(x){x*100}) +
    theme(legend.position = "None",
          aspect.ratio = 1,
          plot.margin = unit(c(0, 0, 0, 0), "lines"),
          axis.title = element_text(size = 6))

p3_surade_roc
```

## fig3 extended plotting
```{r fig3_extended_plotting}
p3_extended <- ggdraw() +
    draw_plot(
        plot_grid(p3_hpylori, p3_tmaritima, p3_surade,
                  labels = c("a", "c", "e"), label_size = 8,
                  ncol = 1, align = "v", rel_heights = c(5, 2.5, 2.5)),
        x = 0, y = 0, width = .8, height = 1) +
    draw_label("Expression Score", x = .03, y = .755, angle = 90, size = 7) + 
    draw_label("Dot Density", x = .03, y = .15, angle = 90, size = 7) + 
    draw_plot(
        plot_grid(p3_hpylori_roc, p3_hpylori_auc, p3_tmaritima_roc,
                  p3_surade_roc, rel_heights = c(1, .8, 1, 1),
                  labels = c("b", "", "d", "f"), label_size = 8,
                  ncol = 1, align = "hv"),
        x = .80, y = 0, width = .17, height = 1)

if (write_out)
    save_plot(filename = "plots/extfig1.pdf", plot = p3_extended,
              base_width = uconv(183, "mm", "in", "Length"), base_height = 5)

p3_extended
```

```{r session}
if (write_out)
    save.image("fig_small_scale.Rdata")
sessionInfo()
```
