---
title: "Data munging and figure preparation"
author: "Shyam Saladi (saladi@caltech.edu)"
date: "09/20/2016"
output: html_document
---

## Load libraries
Load all libraries necessary for subsequent data munging and plotting.
Some defaults are set here to make things easier throughout.
```{r libraries}
# if TRUE, then data will be written to file
write_out = FALSE

library(tibble)
library(dplyr)
library(dplyrExtras)
library(tidyr)

library(caret)
library(pROC)

library(ggplot2)
library(cowplot)
library(corrplot)
library(viridis)
theme_set(theme_cowplot(font_size = 7))

library(foreach)
library(doMC)
registerDoMC(cores = 10)
```

## Convenience functions
Load another file with a number of helper functions to mechanize some tedious
or repetitive tasks.
```{r convenience_fns}
source("preparation-convenience.R")
source("predict.R")
```

## Load SVM weights and calculate correlations
This block loads a csv file of the model's weights and renames certain features
whose original identifiers could be more accurately written by the feature
calculation Python scripts.
```{r load_weights}
# load preprocessing script
# features we use for the ML
load("predict.RData")

# name corrections (see smsaladi/ml-expression#16)
rename_feats <- function(df) {
    rename(df,
           avgCU_first5 = minCU_win5,
           avgCU_first10 = minCU_win10,
           nterm_hydOCT = nterm_hydGES,
           avgRONNTM1_2 = RONNlongestloop,
           RONNlongestloop = avgRONNTM1_2)
}

# load weights
svmweights <- read.csv("daley_only_raw_exp_ml21.full.4.svmweights.csv",
                         header = TRUE) %>%
    mutate(abs_weight = abs(weight),
           rel = weight / mean(weight),
           old_feat = feature) %>%
    # rename features
    mutate_rows(feature == "minCU_win5", feature = "avgCU_first5") %>%
    mutate_rows(feature == "minCU_win10", feature = "avgCU_first10") %>%
    mutate_rows(feature == "nterm_hydGES", feature = "nterm_hydOCT") %>%
    mutate_rows(feature == "RONNlongestloop", feature = "avgRONNTM1_2X") %>%
    mutate_rows(feature == "avgRONNTM1_2", feature = "RONNlongestloop") %>%
    mutate_rows(feature == "avgRONNTM1_2X", feature = "avgRONNTM1_2")
```

## Calculate feature correlations
Calculates the Pearson correlation coefficients between each feature within a
reference dataset, i.e. all forward predicted genomes.
```{r feature_correlation}
# correlation matrix for forward predicted genomes
forward_cor <- forward_features %>%
    preprocess_vars(daley_allstats_exp_xtrans_params) %>%
    rename_feats %>%
    cor(use = "pairwise.complete.obs")

# significance test for correlation matrix
# from corrplot::corplot-intro vignette
forward_cor_mtest <- forward_features %>%
    preprocess_vars(daley_allstats_exp_xtrans_params) %>%
    rename_feats %>%
    cor.mtest

# box/grouping memberships
# adapted from corrplot::corrRect.hclust, the underlying code to generate
# boxes for groupings/orderings
svmweights <- as.dist(1 - forward_cor) %>%
    hclust(method = "complete") %>%
    cutree(k = 11) %>%
    data.frame %>% 
    rownames_to_column %>%
    right_join(svmweights, by = c("rowname" = "feature")) %>%
    rename_(cluster = as.name("."),
            feature = "rowname")

# Add feature groupings back to feature weights dataframe
# with hclust cluster/manual category summary information (i.e. number of
# features that compose the grouping).
svmweights <- svmweights %>%
    group_by(category) %>%
    summarize(man_weight = sum(abs(weight)),
              man_count = n()) %>%
    right_join(svmweights, by = c("category")) %>%
    mutate(man_rank = dense_rank(-man_weight),
           man_alpha = LETTERS[man_rank])

svmweights <- svmweights %>%
    group_by(cluster) %>%
    summarize(hclust_weight = sum(abs(weight)),
              hclust_count = n()) %>%
    right_join(svmweights, by = c("cluster")) %>%
    mutate(hclust_rank = dense_rank(-hclust_weight),
           hclust_roman = as.roman(hclust_rank))
```

## plot summary of weights based on groupings
```{r weight_summary}
p_weight_man_sum <- svmweights %>%
    distinct(category, man_rank, man_alpha, man_weight, man_count) %>% 
    arrange(man_weight) %>%
    mutate(man_label = paste0(man_alpha, ".") %>% paste(category),
           man_label =
               factor(man_label, levels = man_label, ordered = TRUE)) %>%
    ggplot(aes(x = man_label)) +
    geom_bar(aes(y = man_weight), stat = "identity") +
    geom_text(aes(y = man_weight + 0.03, label = man_count), size = 5*5/14) +
    ylab("Total Weight Magnitude") +
    ggtitle("Manually-curated groups") +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    coord_flip() +
    theme(axis.title.y = element_blank())

p_weight_hclust_sum <- svmweights %>%
    distinct(hclust_rank, hclust_roman, hclust_weight, hclust_count) %>% 
    arrange(hclust_weight) %>%
    mutate(hclust_roman = as.character(hclust_roman) %>%
               factor(., levels = ., ordered = TRUE)) %>%
    ggplot(aes(x = hclust_roman)) +
    geom_bar(aes(y = hclust_weight), stat = "identity") +
    geom_text(aes(y = hclust_weight + 0.03, label = hclust_count),
              size = 5*5/14) +
    xlab("Cluster ID") +
    ylab("Total Weight Magnitude") +
    ggtitle("Hierarchical-clustered groups") +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    theme(axis.title.y = element_blank()) +
    coord_flip()

if (!write_out) {
    p_weight_man_sum
    p_weight_hclust_sum
}
```

## plot AUC within NYCOMPS with subsets of features
```{r nycomps_subsets}
calc_auc_feat_subsets <-
    function(use_feat_col, order_fun, order_name,
             ref_feat, ref_eval, weights = svmweights) {
    # transform features
    feat_weights <- setNames(weights$weight, as.character(weights$old_feat))
    trans_feat <- preprocess_vars(ref_feat, daley_allstats_exp_xtrans_params)
    # reorder vector
    feat_weights <- feat_weights[colnames(trans_feat)]
    for (i in seq_along(trans_feat))
        set(trans_feat, j = i, value = trans_feat[[i]] * feat_weights[i])

    trans_feat[is.na(trans_feat)] <- 0

    foreach(thisiter = unique(weights[, use_feat_col] %>% unlist),
            .combine = bind_rows, .multicombine = TRUE) %dopar% {
                # specify those features to use
                # can't seem to work out NSE, so this clumsy way in the meantime
                use_feats <-
                    weights[order_fun(weights[,use_feat_col], thisiter),] %>%
                    select(old_feat) %>%
                    unlist %>% as.character

                # prediction
                preds <- trans_feat[, use_feats, drop = FALSE] %>% rowSums

                # calculate auc
                data.frame(score = ci.auc(response = ref_eval,
                                         predictor = preds,
                                         direction = "<"),
                           ci_labels = c("lower95", "auc", "upper95"),
                           iter = thisiter,
                           feat_count = length(use_feats))
            } %>%
    dcast(iter + feat_count ~ ci_labels, value.var = "score") %>%
        mutate(order_name = order_name,
               use_feat_col = use_feat_col)
}

great_eq <- function(a, b) {
    if (is.na(a)) {
        is.na(b)
    } else {
        a >= b
    }
}
less_eq <- function(a, b) {
    if (is.na(a)) {
        is.na(b)
    } else {
        a <= b
    }
}
eq_fun <- function(a, b) {
    if (is.na(a)) {
        is.na(b)
    } else {
        a == b
    }
}

calc_auc_feat_subsets_wrapper <- function(combos, ...) {
    foreach(row = iter(feat_subset_combos, by = 'row'),
            .combine = bind_rows, .multicombine = TRUE) %do% {
                    if (row$order == 'desc') {
                        calc_auc_feat_subsets(use_feat_col = row$use_feat_col,
                                              order_fun = great_eq,
                                              order_name = row$order,
                                              ...)
                    } else if (row$order == 'incr') {
                        calc_auc_feat_subsets(use_feat_col = row$use_feat_col,
                                              order_fun = less_eq,
                                              order_name = row$order,
                                              ...)
                    } else if (row$order == 'eq') {
                        calc_auc_feat_subsets(use_feat_col = row$use_feat_col,
                                              order_fun = eq_fun,
                                              order_name = row$order,
                                              ...)
                    }
            }
}

feat_subset_combos <- expand.grid(
    use_feat_col = c("abs_weight", "hclust_weight", "man_weight"),
    order_name = c('desc', 'incr', 'eq'))

feat_subset_posmix_auc <- calc_auc_feat_subsets_wrapper(
    combos = feat_subset_combos,
    ref_feat = nycomps_posmix,
    ref_eval = nycomps_posmix$gene_outcome != "neg")

feat_subset_vectors_auc <- calc_auc_feat_subsets_wrapper(
    combos = feat_subset_combos,
    ref_feat = nycomps_vectors,
    ref_eval = nycomps_vectors$plasmid_outcome == 1)
```

```{r}
feat_subset_posmix_auc %>%
    mutate(dataset = "posmix") %>%
    bind_rows(feat_subset_vectors_auc %>% mutate(dataset = "vectors")) %>%
    filter(order_name == "eq", use_feat_col == "abs_weight") %>%
    left_join(svmweights, by = c("iter" = "abs_weight")) %>%
    ggplot(aes(x = iter, y = auc, color = dataset, label = feature)) +
    geom_point() + 
    geom_text(size = 5*5/14) +
    theme(aspect.ratio = 1)
    #facet_wrap(~use_feat_col, ncol = 1, scales = "free_x")
```

```{r panel_indiv_feat_desc}
get_weight <- function(x, match_col, out_col, dataset) {
    foreach(thisx = x, .combine = c) %do% {
        # dplyr way of doing this sucks
        dataset[dataset[, match_col] == thisx, ] %>%
            select_(out_col) %>% unlist %>% as.numeric
    } %>% specify_decimal(k = 2)
}

p4_temp <- feat_subset_posmix_auc %>%
    filter(order_name == "desc", use_feat_col == "abs_weight")

p4_base <- ggplot(p4_temp, aes(feat_count, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    xlab("Number of Features (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(15, 81, 15), 89))  +
    scale_y_continuous(expand = c(0, 0), limits = c(0.46, 0.62),
                       breaks = seq(0.48, 0.62, 0.04),
                       labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

get_indiv_weight <- function(x)
    get_weight(x, match_col = "feat_count", out_col = "iter", dataset = p4_temp)

p4_weights <- 
    ggplot(p4_temp, aes(feat_count, auc)) +
    geom_line() +
    xlab("Feature Magnitude") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(15, 81, 15), 89),
                       labels = get_indiv_weight)

p4_feat_imp_all <- secondary_y_axis(p4_base, p4_weights)
rm(p4_temp, p4_base, p4_weights)
```

## manually curated groups
```{r plot_manually_curated_groups}
p4_temp <- feat_subset_posmix_auc %>%
    filter(order_name == "desc", use_feat_col == "man_weight") %>%
    left_join(svmweights %>% distinct(man_weight, man_rank, man_alpha),
              by = c("iter" = "man_weight"))

p4_base <- ggplot(p4_temp, aes(man_rank, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    geom_text(aes(y = auc + .003, label = man_alpha), size = 5*5/14) +
    xlab("Number of Feature Categories (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(3, 15, 3))) +
    scale_y_continuous(expand = c(0, 0), limits = c(0.46, 0.62),
                       breaks = seq(0.48, 0.62, 0.04),
                       labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

get_man_weight <- function(x)
    get_weight(x, match_col = "man_rank", out_col = "iter", dataset = p4_temp)

p4_weights <- ggplot(p4_temp, aes(man_rank, auc)) +
    geom_line() +
    xlab("Group Magnitude") +
    scale_x_continuous(expand = c(0, 0),
                       breaks = c(1, seq(3, 15, 3)),
                       labels = get_man_weight)

p4_feat_imp_man <- secondary_y_axis(p4_base, p4_weights)
rm(p4_temp, p4_base, p4_weights)

if (!write_out)
    ggdraw(p4_feat_imp_man)
```

## hclust groups
```{r}
p4_temp <- feat_subset_posmix_auc %>%
    filter(order_name == "desc", use_feat_col == "hclust_weight") %>%
    left_join(svmweights %>% distinct(hclust_weight, hclust_rank, hclust_roman),
              by = c("iter" = "hclust_weight"))

p4_base <-
    ggplot(p4_temp, aes(hclust_rank, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    geom_text(aes(y = auc + 0.003,label = as.character(hclust_roman)),
              size = 5*5/14) +
    xlab("Number of Feature Categories (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(2, 12, 2))) +
    scale_y_continuous(expand = c(0, 0), limits = c(0.46, 0.62),
                       breaks = seq(0.48, 0.62, 0.04),
                       labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

get_hclust_weight <- function(x)
    get_weight(x, match_col = "hclust_rank", out_col = "iter", dataset = p4_temp)

p4_weights <- ggplot(p4_temp, aes(hclust_rank, auc)) +
    geom_line() +
    xlab("Group Magnitude") +
    scale_x_continuous(expand = c(0, 0),
                       breaks = c(1, seq(2, 12, 2)),
                       labels = get_hclust_weight)

p4_feat_imp_hclust <- secondary_y_axis(p4_base, p4_weights)
rm(p4_temp, p4_base, p4_weights)

if (!write_out)
    ggdraw(p4_feat_imp_hclust)
```

## Figure importance plot
This block plots the entire figure relating to feature importances.
```{r plot_feature_importances}
p5_weight_importance <- plot_grid(
    plot_grid(p_weight_man_sum, p_weight_hclust_sum,
              ncol = 1, align = "hv",
              labels = letters[1:2], label_size = 8),
    plot_grid(p4_feat_imp_all, p4_feat_imp_man, p4_feat_imp_hclust,
              ncol = 1, align = "hv",
              labels = letters[3:5], label_size = 8),
    nrow = 1) +
    draw_label("Cluster ID", x = 0.13, y = 0.25, angle = 90, size = 7)

if (write_out) {
    save_plot(filename = "plots/p5_weights.pdf", plot = p5_weight_importance,
          base_width = uconv(183, "mm", "in", "Length"),
          base_height = 5.5)
} else {
    p5_weight_importance
}
```

# all weights plotting
```{r fig_feats_all_combos}
feat_subset_posmix_auc %>%
    mutate(dataset = "posmix") %>%
    bind_rows(feat_subset_vectors_auc %>%
                  mutate(dataset = "vectors")) %>%
    mutate(order_name = factor(order_name,
                               levels = c("desc", "incr"),
                               labels = c("Decreasing\nWeight",
                                          "Increasing\nWeight")),
           use_feat_col = factor(use_feat_col,
                                 levels = c("abs_weight", "man_weight",
                                            "hclust_weight"),
                                 labels = c("Individual Features",
                                            "Manually Curated Groups",
                                            "Hierarchially Clustered Groups")),
           dataset = factor(dataset,
                            levels = c("posmix", "vectors"),
                            labels = c("By Gene", "By Expression Trial"))) %>%
    ggplot(aes(feat_count, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95, fill = order_name),
                alpha = 0.5) +
    geom_line(aes(color = order_name)) +
    xlab("Number of Features") +
    ylab("AUC") +
    scale_color_viridis(discrete = TRUE, end = 0.5,
                        guide = guide_legend(title = "Order of Features",
                                             nrow = 1)) +
    scale_fill_viridis(discrete = TRUE, end = 0.5,
                        guide = guide_legend(title = "Order of Features")) +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(0, 81, 15), 89),
                       limits = c(1,89))  +
    scale_y_continuous(expand = c(0, 0), labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"),
          legend.position = c(0.85, 0.07)) +
    facet_grid(dataset~use_feat_col, scales = "free_x")
```

## clustering of entire forward genomes
 `corrplot` is used to generate plot the correlation matrix becuase it more
feature-rich than equvilient ggplot-compatible packages. As a result, the panel
is saved and stiched together with the others after-the-fact using Adobe
Illustrator.
```{r plot_big_clustering}
if (write_out)
    pdf("plots/forward_corr.pdf", width = 10, height = 10)
# `corrplot` returns the correlation matrix re-ordered after clustering as
# plotted. Save this for the weights plot (chunk below).
forward_cor_clust <-
    corrplot(forward_cor_clust, col = viridis(300),
             method = "square", plotCI = "n",
             order = "hclust", hclust.method = "complete", addrect = 11,
             p.mat = forward_cor_mtest[[1]], sig.level = 0.05, insig = "blank",
             lowCI.mat = forward_cor_mtest[[2]],
             uppCI.mat = forward_cor_mtest[[3]],
             tl.cex = .5, tl.col = "black", tl.srt = 45)
if (write_out)
    dev.off()
```

## Plot weights
This block plots the weights for the SVM model trained in the order of the 
hclust clustering of the correlation matrix (plotted directly above). 
```{r plot_weights}
feat_order <- dimnames(forward_cor_clust)[[1]] %>% rev
p_svmweights <- svmweights %>%
    # these were not calculated across the full forward genomes set
    filter(!(feature %in% c("avgRNAss", "minRNAss", "maxRNAss",
                            "q75RNAss", "q25RNAss"))) %>%
    mutate(feature =
               factor(feature, labels = feat_order, levels = feat_order)) %>%
    na.omit %>%
    ggplot() +
    geom_bar(aes(x = feature, y = weight), stat = "identity") +
    ylab(expression("SVM" ^ "rank"*~"score")) +
    coord_flip() +
    theme(axis.title.y = element_blank())

if (write_out) {
    save_plot("plots/svmweights.pdf", p_svmweights,
              base_height = 10, base_width = 3)
} else {
    p_svmweights
}
```

```{r session}
if (write_out)
    save.image("fig_features.Rdata")
sessionInfo()
```
