---
title: "Data munging and figure preparation"
author: "Shyam Saladi (saladi@caltech.edu)"
date: "09/20/2016"
output: html_document
---

## Load libraries
Load all libraries necessary for subsequent data munging and plotting.
Some defaults are set here to make things easier throughout.
```{r libraries}
# if TRUE, then data will be written to file
write_out = FALSE

library(tibble)
library(dplyr)
library(dplyrExtras)
library(tidyr)

library(caret)

library(ggplot2)
library(cowplot)
library(corrplot)
library(viridis)
theme_set(theme_cowplot(font_size = 7))

library(foreach)
library(doMC)
registerDoMC(cores = 10)
```

## Convenience functions
Load another file with a number of helper functions to mechanize some tedious
or repetitive tasks.
```{r}
source("preparation-convenience.R")
source("predict.R")
```

# plot scores of features along with correlations within
# ecoli training, nycomps, and forward genomes datasets
```{r feature_plot}
# load preprocessing script
# features we use for the ML
load("predict.RData")

# name corrections (see smsaladi/ml-expression#16)
rename_feats <- function(df) {
    rename(df,
           avgCU_first5 = minCU_win5,
           avgCU_first10 = minCU_win10,
           nterm_hydOCT = nterm_hydGES,
           avgRONNTM1_2 = RONNlongestloop,
           RONNlongestloop = avgRONNTM1_2)
}

# load weights 
svmweights <- read.csv("daley_only_raw_exp_ml21.full.4.svmweights.csv",
                         header = TRUE) %>%
    mutate(rel = weight / mean(weight),
           old_feat = feature) %>%
    # rename features
    mutate_rows(feature == "minCU_win5", feature = "avgCU_first5") %>%
    mutate_rows(feature == "minCU_win10", feature = "avgCU_first10") %>%
    mutate_rows(feature == "nterm_hydGES", feature = "nterm_hydOCT") %>%
    mutate_rows(feature == "RONNlongestloop", feature = "avgRONNTM1_2X") %>%
    mutate_rows(feature == "avgRONNTM1_2", feature = "RONNlongestloop") %>%
    mutate_rows(feature == "avgRONNTM1_2X", feature = "avgRONNTM1_2")

# correlation matrix for forward predicted genomes
forward_cor <- forward_features %>%
    preprocess_vars(daley_allstats_exp_xtrans_params) %>%
    rename_feats %>%
    cor(use = "pairwise.complete.obs")

# significance test for correlation matrix
forward_cor_mtest <- forward_features %>%
    preprocess_vars(daley_allstats_exp_xtrans_params) %>%
    rename_feats %>%
    cor.mtest

# box/grouping memberships
# from corrplot::corrRect.hclust
svmweights <- hclust(as.dist(1 - forward_cor), method = "complete") %>%
    cutree(k = 11) %>%
    data.frame %>% 
    rownames_to_column %>%
    right_join(svmweights, by = c("rowname" = "feature")) %>%
    rename_(cluster = as.name("."),
            feature = "rowname")

svmweights <- svmweights %>%
    group_by(category) %>%
    summarize(man_weight = sum(abs(weight)),
              man_count = n()) %>%
    right_join(svmweights, by = c("category"))

svmweights <- svmweights %>%
    group_by(cluster) %>%
    summarize(hclust_weight = sum(abs(weight)),
              hclust_count = n()) %>%
    right_join(svmweights, by = c("cluster"))
```

## plot summary of weights based on groupings
```{r weight_summary}
man_feat_order <- svmweights %>%
    arrange(man_weight) %>%
    distinct(category) %>%
    unlist %>% as.character

p_weight_man_sum <- svmweights %>%
    distinct(category, man_weight, man_count) %>%
    mutate(category = factor(category, levels = man_feat_order)) %>%
    ggplot(aes(x = category)) +
    geom_bar(aes(y = man_weight), stat = "identity") +
    geom_text(aes(y = man_weight + 0.03, label = man_count), size = 5*5/14) +
    ylab("Total Weight Magnitude") +
    ggtitle("Manually-curated groups") +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    coord_flip() +
    theme(axis.title.y = element_blank())

hclust_feat_order <- svmweights %>%
    arrange(hclust_weight) %>%
    distinct(cluster) %>%
    unlist %>% as.character

p_weight_hclust_sum <- svmweights %>%
    distinct(cluster, hclust_weight, hclust_count) %>%
    mutate(cluster = factor(cluster, levels = hclust_feat_order)) %>%
    na.omit %>%
    ggplot(aes(x = cluster)) +
    geom_bar(aes(y = hclust_weight), stat = "identity") +
    geom_text(aes(y = hclust_weight + 0.03, label = hclust_count),
              size = 5*5/14) +
    xlab("Cluster ID") +
    ylab("Total Weight Magnitude") +
    ggtitle("Hierarchical-clustered groups") +
    scale_x_discrete(expand = c(0, 0)) +
    scale_y_continuous(expand = c(0, 0)) +
    coord_flip()

if (!write_out) {
    p_weight_man_sum
    p_weight_hclust_sum
}
```

## plot AUC within NYCOMPS with subsets of features
```{r nycomps_subsets}
svmweights <- svmweights %>% arrange(desc(abs(weight)))

nycomps_auc_feat_all_subsets <-
    foreach(maxidx = seq.int(1, nrow(svmweights)),
            .combine = bind_rows, .multicombine = TRUE) %dopar% {
    # specify those features to use
    use_feats <- as.character(svmweights[seq(1, maxidx),]$old_feat)
    
    # run prediction
    preds <- prediction_fn(
        test_data = nycomps_posmix[, use_feats, drop = FALSE],
        xtrans = daley_allstats_exp_xtrans_params,
        model_dat_fn = "daley_only_raw_exp_ml21.full.4.model",
        classify_cmd = "/ul/saladi/apps/svm_rank/svm_rank_classify")
    
    # calculate auc
    data.frame(ml21 = ci.auc(response = nycomps_posmix$gene_outcome != "neg",
                             predictor = preds,
                             direction = "<"),
               ci_labels = c("lower95", "auc", "upper95"),
               thisidx = maxidx,
               svmweights[maxidx,])
} %>%
    dcast(thisidx + weight + rel ~ ci_labels, value.var = "ml21")

nycomps_auc_feat_hclust <-
    foreach(thisweight =
                unique(svmweights$hclust_weight) %>% sort(decreasing = TRUE),
            .combine = bind_rows, .multicombine = TRUE) %dopar% {
    # specify those features to use
    use_feats <- svmweights %>%
        filter(hclust_weight >= thisweight) %>%
        select(old_feat) %>%
        unlist %>% as.character
    
    # run prediction
    preds <- prediction_fn(
        test_data = nycomps_posmix[, use_feats, drop = FALSE],
        xtrans = daley_allstats_exp_xtrans_params,
        model_dat_fn = "daley_only_raw_exp_ml21.full.4.model",
        classify_cmd = "/ul/saladi/apps/svm_rank/svm_rank_classify")
    
    # calculate auc
    data.frame(ml21 = ci.auc(response = nycomps_posmix$gene_outcome != "neg",
                             predictor = preds,
                             direction = "<"),
               ci_labels = c("lower95", "auc", "upper95"),
               hclust_weight = thisweight)
} %>%
    dcast(hclust_weight ~ ci_labels, value.var = "ml21") %>%
    left_join(svmweights %>% distinct(cluster, hclust_count, hclust_weight),
              by = c("hclust_weight"))

nycomps_auc_feat_man <-
    foreach(thisweight =
                unique(svmweights$man_weight) %>% sort(decreasing = TRUE),
            .combine = bind_rows, .multicombine = TRUE) %dopar% {
    # specify those features to use
    use_feats <- svmweights %>%
        filter(man_weight >= thisweight) %>%
        select(old_feat) %>%
        unlist %>% as.character
    
    # run prediction
    preds <- prediction_fn(
        test_data = nycomps_posmix[, use_feats, drop = FALSE],
        xtrans = daley_allstats_exp_xtrans_params,
        model_dat_fn = "daley_only_raw_exp_ml21.full.4.model",
        classify_cmd = "/ul/saladi/apps/svm_rank/svm_rank_classify")
    
    # calculate auc
    data.frame(ml21 = ci.auc(response = nycomps_posmix$gene_outcome != "neg",
                             predictor = preds,
                             direction = "<"),
               ci_labels = c("lower95", "auc", "upper95"),
               man_weight = thisweight)
} %>%
    dcast(man_weight ~ ci_labels, value.var = "ml21") %>%
    left_join(svmweights %>% distinct(category, man_count, man_weight),
              by = c("man_weight"))
```

# all weights plotting
```{r nycomps_feat_weight_plotting}
p4_base <- ggplot(nycomps_auc_feat_all_subsets, aes(thisidx, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    xlab("Number of Features (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(0, 81, 15), 89))  +
    scale_y_continuous(expand = c(0, 0), labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

p4_weights <- ggplot(nycomps_auc_feat_all_subsets, aes(thisidx, auc)) +
    geom_line() +
    xlab("Feature Magnitude") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, seq(0, 81, 15), 89),
                       labels = function(x)
                           nycomps_auc_feat_subsets_posmix[x, "weight"] %>%
                           abs %>% specify_decimal(k = 2))

p4_feat_imp_all <- secondary_y_axis(p4_base, p4_weights)
rm(p4_base, p4_weights)
```

## manually curated groups
```{r}
nycomps_auc_feat_man <- nycomps_auc_feat_man %>%
    mutate(rank = rank(-man_weight))

p4_base <- ggplot(nycomps_auc_feat_man, aes(rank, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    geom_text(aes(label = category), size = 5*5/14) +
    xlab("Number of Feature Categories (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, 5, 10, 15)) +
    scale_y_continuous(expand = c(0, 0), labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

get_man_weight <- function(x) {
    foreach(thisx = x, .combine = c) %do% {
        filter(nycomps_auc_feat_man, rank == thisx) %>%
            select(man_weight) %>% unlist %>% as.numeric
    } %>% specify_decimal(k = 2)
}

p4_weights <- ggplot(nycomps_auc_feat_man, aes(rank, auc)) +
    geom_line() +
    xlab("Group Magnitude") +
    scale_x_continuous(expand = c(0, 0),
                       breaks = c(1, 5, 10, 15),
                       labels = get_man_weight)

p4_feat_imp_man <- secondary_y_axis(p4_base, p4_weights)
rm(p4_base, p4_weights)

if (!write_out)
    ggdraw(p4_feat_imp_man)
```

## hclust groups
```{r}
nycomps_auc_feat_hclust <- nycomps_auc_feat_hclust %>%
    mutate(rank = rank(-hclust_weight))

p4_base <- ggplot(nycomps_auc_feat_hclust %>% na.omit, aes(rank, auc)) +
    geom_hline(yintercept = 0.5, linetype = "longdash", color = "gray") + 
    geom_ribbon(aes(ymin = lower95, ymax = upper95), fill = "#A4A5A5") +
    geom_line() +
    geom_text(aes(label = cluster), size = 5*5/14) +
    xlab("Number of Feature Categories (decreasing in weight)") +
    ylab("AUC") +
    scale_x_continuous(expand = c(0, 0), breaks = c(1, 3, 6, 9, 12)) +
    scale_y_continuous(expand = c(0, 0), labels = function(x) x*100) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "in"))

get_hclust_weight <- function(x) {
    foreach(thisx = x, .combine = c) %do% {
        filter(nycomps_auc_feat_hclust, rank == thisx) %>%
            select(hclust_weight) %>% unlist %>% as.numeric
    } %>% specify_decimal(k = 2)
}

p4_weights <- ggplot(nycomps_auc_feat_hclust %>% na.omit, aes(rank, auc)) +
    geom_line() +
    xlab("Group Magnitude") +
    scale_x_continuous(expand = c(0, 0),
                       breaks = c(1, 3, 6, 9, 12),
                       labels = get_hclust_weight)

p4_feat_imp_hclust <- secondary_y_axis(p4_base, p4_weights)
rm(p4_base, p4_weights)

if (!write_out)
    ggdraw(p4_feat_imp_hclust)
```

## compile weights figure
```{r}
p5_weights <- plot_grid(
    plot_grid(p_weight_man_sum, p_weight_hclust_sum,
              ncol = 1, align = "hv",
              labels = letters[1:2], label_size = 8),
    plot_grid(p4_feat_imp_all, p4_feat_imp_man, p4_feat_imp_hclust,
              ncol = 1, align = "hv",
              labels = letters[3:5], label_size = 8),
    nrow = 1)

if (write_out) {
    save_plot(filename = "plots/p5_weights.pdf", plot = p5_weights,
          base_width = uconv(183, "mm", "in", "Length"),
          base_height = 5.5)
} else {
    p5_weights
}
```

## clustering of entire forward genomes
```{r big_clustering}
if (write_out)
    pdf("plots/forward_corr.pdf", width = 10, height = 10)
forward_cor_clust <-
    corrplot(forward_cor_clust, col = viridis(300),
             method = "square", plotCI = "n",
             order = "hclust", hclust.method = "complete", addrect = 11,
             p.mat = forward_cor_mtest[[1]], sig.level = 0.05, insig = "blank",
             lowCI.mat = forward_cor_mtest[[2]],
             uppCI.mat = forward_cor_mtest[[3]],
             tl.cex = .5, tl.col = "black", tl.srt = 45)
if (write_out)
    dev.off()
```

## plot weights in the order of the clustering
```{r svmweights_cluster_order}
feat_order <- dimnames(forward_cor_clust)[[1]] %>% rev
p_svmweights <- svmweights %>%
    # these are missing from the full forward set
    filter(!(feature %in% c("avgRNAss", "minRNAss", "maxRNAss",
                            "q75RNAss", "q25RNAss"))) %>%
    mutate(feature =
               factor(feature, labels = feat_order, levels = feat_order)) %>%
    na.omit %>%
    ggplot() +
    geom_bar(aes(x = feature, y = weight), stat = "identity") +
    ylab(expression("SVM" ^ "rank"*~"score")) +
    coord_flip() +
    theme(axis.title.y = element_blank())

if (write_out) {
    save_plot("plots/svmweights.pdf", p_svmweights, base_height = 10, base_width = 5)
} else {
    p_svmweights
}
```

```{r session}
if (write_out)
    save.image("fig_features.Rdata")
sessionInfo()
```
